Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\

### 2024-10-08 main.tex
Chapter V.A - Experimentsal Evaluation - Setup; still WIP

\newcommand{\heftmm}{\algo{HEFTM\_MM}\xspace}
\newcommand{\heftbl}{\algo{HEFTM\_BL}\xspace}
\newcommand{\heftblc}{\algo{HEFTM\_BLC}\xspace}

@@

        %Scheduling scientific workflows is important. (\skug{can't imagine a good first sentence})
        Scientific workflows are often represented as directed acyclic graphs (DAGs),
        where vertices correspond to tasks and edges represent the dependencies between them.
        Typically, each task requires a
        certain amount of memory to be executed and needs to communicate data to its successor tasks.
        The goal is generally to execute the workflow  as fast as possible (i.e., to minimize its makespan), 
        while satisfying the memory constraints.  
        Hence, we investigate the memory-aware scheduling of DAG-shaped workflows on 
        heterogeneous platforms, where each processor can have a different speed and a different memory size.
        We propose a variant of HEFT (Heterogeneous Earliest Finish Time) that (in contrast to the original) accounts for memory and 
        includes eviction strategies for cases when it might be beneficial to remove some data from memory
	in order to have enough memory to execute other tasks.
%
	Furthermore, while HEFT assumes perfect knowledge of the execution time and memory usage
	of each task, the actual values might differ upon execution. Thus, we propose an adaptive
	scheduling strategy, where a schedule is recomputed when there has been a significant variation in terms
	of execution time or memory. 
%	
	The scheduler has been closely integrated with a runtime system, allowing us to perform a thorough
	experimental evaluation on real-world workflows. The runtime system warns the scheduler when 
	the task parameters have changed, and a schedule can be recomputed on the fly. The memory-aware
	strategy allows us to schedule task graphs that would run out of memory with a state-of-the-art
	scheduler, and the adaptive setting allows us to significantly reduce the makespan.

%                when tentatively assigning tasks, in order to 
%        Its first step is to compute the weights of the tasks.
%        We suggest three variants: bottom levels as weights, bottom levels with impact of incoming edge weight,
%        and weights along the optimal memory traversal.
%        In the second step, we try assigning each task to each processors and execute the assignment that
%        is feasible with regard to memory size and gives the earliest finishing time to the task.
%        Sometimes, data corresponding to edge weights that is stored in the memory needs to be evicted in order to
%        assign a task to a processor.
%        We suggest two eviction strategies - largest files first and smallest files first.

%        Our experimental evaluation on real-world workflows and simulated (\skug{generated? They are generated from real-world wfs})
%        ones with real task and edge weights with up to 30,000 tasks shows that
%        respecting memory constraints only costs $11\%$ of runtime in comparison to a non memory-aware baseline.
%        Calculating task weights with the impact of memory gives a $x\%$ better makespans on a normal and $y\%$ better makespans
%        on small one.
%        Calculating task weights along the optimal memory traversal gives on average $z\%$ worse makespans, but improves
%        average memory utlization by $t\%$.

@@

%    \begin{IEEEkeywords}
%        DAG, Heterogeneous platform, Adaptive scheduling, Memory constraint.
%    \end{IEEEkeywords}

    \section{Introduction} %: \skug{Full: 0, Polished: 0}}

%    \skug{Fullness score (how much text is available): 0-none to 5 - everything.
%
%    Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
%    }
%    TODO: Insert abstract

@@

    All algorithms are implemented in C++ and compiled with g++ (v.11.2.0).
    The experiments are managed by simexpal~\cite{DBLP:journals/algorithms/AngrimanGLMNPT19} and executed on workstations with 192 GB RAM and 2x 12-Core Intel Xeon 6126 @3.2 GHz
    and CentOS 8 as OS.
    Code, input data, and experiment scripts are available for review under \url{TODO:insert link}.

@@

    \subsubsection{Workflow instances}
    \skug{Todo: describe the workflows the same asin the abstract}
    The input data for the experiments consists of two sets of workflows: real-world workflows
    obtained from~\cite{ewels2020nf} and workflows obtained by simulating real-world workflows
    with the WFGen generator~\cite{COLEMAN202216}.
    First, we discuss how the graph topology is generated,
    then we focus on the weights associated to tasks and edges.

    \paragraph{Workflow graphs}
    For real-world workflows, their nextflow definition (see~\cite{ewels2020nf}) was downloaded from the
    respective repository and transformed into .dot format using the nextflow option ``-with-dag''.
    The resulting DAG contains many pseudo-tasks that are only internal representations in nextflow
    (and not actual tasks); that is why we removed them.

    For the simulated workflows, the graph is produced by the WFGen generator, based on a {\em model workflow} and
    the desired number of tasks.
    The model workflows are described on the WFGen website, and we used the following ones:
    1000Genome, BLAST, BWA, Epigenomics, Montage, Seismology, and SoyKB.
    Other models could not be generated without errors.
%
    As number of tasks, we use: 200, \numprint{1000}, \numprint{2000}, \numprint{4000}, \numprint{8000}, \numprint{10000},
    \numprint{15000}, \numprint{18000}, \numprint{20000}, \numprint{25000}, \numprint{30000}.
    We divide the workflows into three groups by size: small ones with up to \numprint{8000} tasks, middle
    ones with \numprint{10000} to \numprint{18000} tasks, and big ones with \numprint{20000} to \numprint{30000} tasks.
    Note that for some workflow models, such as SoyKB or Montage, only a subset of the workflow sizes could be generated,
    because the workflow generator either took a disproportionally long time or yielded errors.

    Overall, this yields four workflow types, denoted by real, small, mid, and big.

    \paragraph{Generation of task and edge weights}
    For the real-world workflows, we use historical data files provided by Bader~\etal~\cite{lotaru}.
    The columns in these files are measured Linux PS stats, acquired during an execution of a nextflow workflow.
    Each row corresponds to an execution of one task on one cluster node.
    Since the operating system cannot distinguish between (a) the RAM the task uses for itself and (b) the RAM it uses
    to store files that were sent or received from other tasks, the values in the historical data are total memory requirements (input/output files plus memory consumption of the computation).
    In a similar manner, the historical data provided by~\cite{lotaru} do not store the actual weights of edges between tasks, but only the overall
    size of files that the task sends to all its children.
    For each task, historical data can contain multiple values, obtained from the runs with different input sizes.
    To avoid underestimation, we take the maximum among all values from different runs on the same cluster node.
    Not all tasks have historical runtime data stored in the tables.
    In fact, for two workflows, Bader~\etal do not provide data for more than 50\% of the tasks.
    For two more, around 40\% of tasks have no historical runtime data stored.
    Hence, in the absence of historical data about a task, we give it a weight of~1.
%
    Because the historical data contains absolute measured values and the cluster node information is a relative value,
    we normalize all values extracted from the historical data by the smallest one.
    This way, task memory weights fit into the memory of the cluster nodes.
    Additionally, this way, the tasks without historical data receive less insignificant values compared to tasks with historical data.

    For the simulated workflows, we generate random execution times and memory weights for each task as well as
    edge weights for each precedence constraint.
    We generate uniformly distributed values between $1$ and $10$ for edge weights,
    $1$ and $1000$ for the workloads,
    and $1$ and $192$ for memory weights.
    When doing so, we try to mimic the weights observed in the historical data, hence \eg the low lower bounds for the
    workloads.

    \subsubsection{Target computing systems}
    \skug{Check that the description of speeds and memories is plausible and consistent with experimental setup. esp check for 19200 -> 192 GB.}
    To fully benefit from the historical data, the  {\em default} experimental environment
    that we consider is a cluster based on the same six
    kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
    We set the number of each kind of node to six, thus having 36 processors in total. % the whole cluster.

    Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as (name, speed, memory):
    ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines; ($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and large memory
    (see Table~\ref{tab:procs} in Appendix~\ref{sub:appx-cluster-config}~\cite{daghetpart_full_version}).
    Note that, by nature,
    memory sizes are normalized values, relative to each other, too.
    Therefore, we additionally normalize memory weights of real-world workflows
    to the maximum size of 192 to make sure they fit.
    For simulated workflows, we increase memory sizes proportionally until the task
    with the biggest memory requirement still
    has a processor it could be executed on.

    To test various settings, we also vary the cluster configuration and consider
    variants of the default cluster presented above:\\
    $\bullet$ {\em Small} and {\em large} clusters. While the default cluster consists of 36 nodes (6 of each kind), we
    consider a small cluster with three processors of each kind ($18$ processors in total), and a
    large cluster with ten processors of each kind ($60$ processors in total). \skug{leaving the big cluster inside for
    the case we do want to include it}\\
    $\bullet$ {\em Fat-and-thin} cluster. In this cluster, the fastest processors have the smallest amount of memory (thin),
    while the slowest processors have the largest memories (fat).
    The overall amount of memory in the cluster, as well as the overall speed of all processors remain the same as in the
            {\em normal} cluster.
%\end{itemize}

    \subsection{Results \skug{Full: 3, Polished: 2}}
    \subsubsection{Static experiments}
    \paragraph{Normal cluster}

>>>>

\newcommand{\heft}{\algo{HEFT}\xspace}
\newcommand{\heftmm}{\algo{HEFTM\_MM}\xspace}
\newcommand{\heftbl}{\algo{HEFTM\_BL}\xspace}
\newcommand{\heftblc}{\algo{HEFTM\_BLC}\xspace}

@@

        Scheduling scientific workflows is important. (\skug{can't imagine a good first sentence})
        They are often represented as directed acyclic graphs (DAGs),
        where vertices correspond to tasks and edges represent the dependencies between them.
       % Since these graphs are often large in both the number of tasks and their resource requirements, it is important
       % to schedule them efficiently on parallel or distributed compute systems.
        Typically, each task requires a
        certain amount of memory to be executed and needs to communicate data to its successor tasks.
        The goal is thus to execute the workflow  as fast as possible (i.e., to minimize its makespan) while satisfying the memory constraints.  %and using parallelism to speed up the execution.

        Hence, we investigate adaptive memory-aware scheduling  of DAG-shaped workflows onto heterogeneous platforms where
        each processor can have a different speed and a different memory size.
        We propose a two-step HEFT-inspired heuristic.
        Its first step is to compute the weights of the tasks.
        We suggest three variants: bottom levels as weights, bottom levels with impact of incoming edge weight,
        and weights along the optimal memory traversal.
        In the second step, we try assigning each task to each processors and execute the assignment that
        is feasible with regard to memory size and gives the earliest finishing time to the task.
        Sometimes, data corresponding to edge weights that is stored in the memory needs to be evicted in order to
        assign a task to a processor.
        We suggest two eviction strategies - largest files first and smallest files first.

        Our experimental evaluation on real-world workflows and simulated (\skug{generated? They are generated from real-world wfs})
        ones with real task and edge weights with up to 30,000 tasks shows that
        respecting memory constraints only costs $11\%$ of runtime in comparison to a non memory-aware baseline.
        Calculating task weights with the impact of memory gives a $x\%$ better makespans on a normal and $y\%$ better makespans
        on small one.
        Calculating task weights along the optimal memory traversal gives on average $z\%$ worse makespans, but improves
        average memory utlization by $t\%$.

@@

    \begin{IEEEkeywords}
        Scheduling, Adaptive Scheduling, DAG
    \end{IEEEkeywords}

    \section{Introduction: \skug{Full: 0, Polished: 0}}

    \skug{Fullness score (how much text is available): 0-none to 5 - everything.

    Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
    }
    TODO: Insert abstract

@@

    All algorithms are implemented in C++ and compiled with g++ (v.8.5.0).
    The experiments are executed on workstations with 192 GB RAM and 2x 12-Core Intel Xeon 6126 @3.2 GHz
    and CentOS 8 as OS.
    Code, input data, and experiment scripts are available for review under \skug{\url{insert link}}.

@@

    \subsubsection{Workflow instances}
    \skug{Todo: describe the workflows the same asin the abstract}
    We run our experiments on a dataqset comprised of real-world workflows: ones used in~\cite{lotaru} (atacseq, bacass, chipseq,
    eager and methylseq), and ones obtained by increasing their size using the WFGen generator~\cite{COLEMAN202216}.

    \paragraph{Workflow graphs}
    For the five real-world workflows, their nextflow definition (see~\cite{ewels2020nf}) was downloaded from the
    respective repository and transformed into .dot format using the nextflow option ``-with-dag''.
    The resulting DAG contains many pseudo-tasks that are only internal representations in nextflow
    (and not actual tasks); that is why we removed them.

    For the size-increased workflows, the graph is produced by the WFGen generator, based on a {\em model workflow} and
    the desired number of tasks.
    We used the real-world workflows as models, except for bacass that lead to errors in the generator.
%
    As number of tasks, we use: 200, \numprint{1000}, \numprint{2000}, \numprint{4000}, \numprint{8000}, \numprint{10000},
    \numprint{15000}, \numprint{18000}, \numprint{20000}, \numprint{25000}, \numprint{30000}.
    We divide the workflows into three groups by size: small ones with up to \numprint{8000} tasks, middle
    ones with \numprint{10000} to \numprint{18000} tasks, and big ones with \numprint{20000} to \numprint{30000} tasks.

    \skug{Im still considerin adding category "tiny" with original workflows and those with 200 tasks. this needs to be double-checked after}
    Overall, this yields four workflow types, denoted by tiny, small, mid, and big.

    \paragraph{Task and edge weights}
    For the real-world workflows, we use historical data files provided by Bader~\etal~\cite{lotaru}.
    The columns in these files are measured Linux PS stats, acquired during an execution of a nextflow workflow.
    Each row corresponds to an execution of one task on one cluster node.
    Since the operating system cannot distinguish between (a) the RAM the task uses for itself and (b) the RAM it uses
    to store files that were sent or received from other tasks, the values in the historical data are total memory requirements (input/output files plus memory consumption of the computation).
    In a similar manner, the historical data provided by~\cite{lotaru} do not store the actual weights of edges between tasks, but only the overall
    size of files that the task sends to all its children.

    For each task, historical data can contain multiple values, obtained from the runs with different input sizes.
    The same workflow can require different memory capacity and take different time to execute
    depending on the size of its input.
    We simulate these various runs by obtaining values corresponding to each input size.
    For each of the four families, there are five input sizes, so we run each workflow in five variants corresponding to these inputs

    Not all tasks have historical runtime data stored in the tables.
    In fact, for two workflows, Bader~\etal do not provide data for more than 50\% of the tasks.
    For two more, around 40\% of tasks have no historical runtime data stored.
    Hence, in the absence of historical data about a task, we give it fixed weights.
    We give it an execution time of $1$, a memory requirement of $50 MB$, and files written and received of $1KB$.
    These values align with the findings of~\cite{lotaru} about small tasks.
%
    \skug{insert more abt how the runtime system hanfles data reads}

    \subsubsection{Target computing systems}
    \skug{Check that the description of speeds and memories is plausible and consistent with experimental setup. esp check for 19200 -> 192 GB.}

    To fully benefit from the historical data, the  {\em default} experimental environment
    that we consider is a cluster based on the same six
    kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
    We set the number of each kind of node to \skug{$24$ - check if this still holds }, thus having 144 processors in total. % the whole cluster.

    Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as (name, speed, memory):
    ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines; ($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and large memory
    (see Table~\ref{tab:procs} in Appendix~\ref{sub:appx-cluster-config}~\cite{daghetpart_full_version}).

    To test various settings, we also vary the cluster configuration and consider
    variants of the default cluster presented above:\\
    $\bullet$ {\em tiny}, {\em small} and {\em large} clusters. While the default cluster consists of 144 nodes (24 of each kind), we
    consider a small cluster with $18$ processors of each kind ($108$ processors in total), a tiny cluster with $72$ nodes and a
    large cluster with $36$ processors of each kind ($216$ processors in total). \skug{check for cluster sizes and naming}\\

    \subsection{Results \skug{Full: 3, Polished: 2}}
    \subsubsection{Static experiments}
    \paragraph{Small cluster}

   On the small cluster, \heft could produce valid assignments in $14$ experiments out of $290$ ($4.8\%$).
    The successful schedules were achieved almost exclusively on the tiny workflows and one one 200-task workflow.
    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$, while \heftmm could schedule
    $94.9\%$ of workflows, thus having the smallest failure rate of $5\%$.

    \paragraph{Normal cluster}
